Spin canonical divisors of log stacky curves

### 2015-06-25 stacky_write_up/spin_cring.tex
removed minimality and changed admissibility notation to a tripple from a pair

Let $R'$ be the canonical ring associated to $\sx'$. Define $(\sx',J)$
to be {\bf admissible} if the $R'$ admits a presentation
\todo{We should incorporate dependence on $L$ into the definition of admissible.}
\begin{align*}
	R' \cong \left( k[x_1, \ldots, x_m] \otimes k[y_{i, e_i'}]_{i \in J} \right)/I'
\end{align*}

@@

\begin{lem}
\label{lem:admissible_subset}
If $(\sx', J)$ is admissible and $W \subset J$ is any subset,
then $(\sx', W)$ is also admissible.
\end{lem}

@@

\begin{thm}
\label{thm:ram_order_ind}
Suppose $(\sx', \Delta, \halfcan')$ is a log spin curve with coarse
space $X'$ and signature $\sigma = (g; e_1', \ldots, e_r'; \delta)$ and either 
\begin{enumerate}
	\item  $(\sx',J)$ is admissible with generators $x_j, y_{i, e_i'} \in R' = R_{L'}$ with $i \in J$ as in Definition ~\ref{defn:admissible}
	\item or $\sigma$ is one of the signatures described in Remark ~\ref{rem:three-cases}, with $y_{1, 3} = y_{2,3} = \ldots, y_{r,3}$ be a rational section of $\sco(\halfcan_X)$ so that $\ord_{P_i}^{\halfcan_X'}(y_{1,3}) = 1$ at $P_i$ for all $i = 1,\ldots, r$.
\end{enumerate}
Let
$(\sx, \Delta, \halfcan)$ be another log spin curve
with coarse space $X = X'$ and signature $(g; e_1, \ldots, e_r;
\delta)$ such that $e_i = e_i' + 2$ for all $i \in J$ and
$e_j = e_j'$ for $j \notin J$. Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i, e_i} \in
		H^0(\sx, e_i(K_\sx))$ so that
		\begin{align*}
			-\ord_{Q_i}
^{\halfcan_X'}(y_{i, e_i}) = \frac{e_i - 1}{2}
		\end{align*}
		and
		\begin{align*}
			\frac{-\ord_{Q_j}
^{\halfcan_X'}(y_{i, e_i})}{\deg (y_{i, e_i})} \leq 
\subhalf{
			e_j'} - \frac{1}{\deg(y_{i, e_i})e_j'}
		\end{align*}
		for all $j \in J$ with $j \neq i$.
	\item[(b)] If $(\sx',\halfcan',J)$ is admissible, then $y_{i, e_i'}^a y_{i, e_i}^b$ with $a \geq 
0,
		b > 0$ span $R$ over $R'$ and the elements $y_{i, e_i}$ minimally
		generate $R$ over $R'$.
	\item[(c)] If $(\sx',\halfcan',J)$ is admissible, equip $k[y] = k[y_{i, e_i}]_{i \in J}$ and $k[x]$ with
		any graded monomial order and $k[y, x] = k[y] \otimes k[x]$ with
		block order. Let $R = k[y, x]/I.$ Then,
		\begin{align*}
			in_\prec(I) = in_\prec(I')k[x, y] + \langle y_{i, e_i}z \mid z 
			\neq y_{i, e_i}, y_{i, e_i'} \rangle
		\end{align*}
		as $z$ ranges over all generators of $R$.
	\item[(d)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}

\begin{proof}
We will only prove this in the case (1), that $(\sx', L', J)$ is admissible, as the case (2) that $\sigma$ is as in Remark ~\ref{rem:three-cases} is completely analogous.

@@

Finally, to check (d), we will first show that $(\sx, J)$ is
admissible. We know (Ad-i) is satisfied by part (b), taking the $y_{
i, e_i}$ as the generators in degree $e_i$. Next, (Ad-ii) is
strictly monotonic in the $e_i$ and hence also holds for $(\sx, J)$.
Finally, if (Ad-iii) holds for $e$ then it holds for $e + 2$ by
definition. This is where we use that (Ad-iii) holds for $d > 0$
and not just for $d = 0$.

Finally, if $(\sx , J)$ is admissible then so is $(\sx, W)$ for any
$W \subset J$ by Lemma ~\ref{lem:admissible_subset}.
\end{proof}

\begin{cor}
\label{cor:raise-stacky-order}
Suppose $(\sx',L',J')$ is admissible with signature $\sigma' = (e_1',\ldots,e_r')$ or $\sigma$ satisfies one of the conditions of Remark ~\ref{rem:three cases}. Let $J' = \{t,t+1,\ldots,r\}$ and $e_1' \leq e_2' \leq \cdots \leq e_t' = e_{t+1}' =\cdots = e_r'$
so that $(\sx',\Delta, L)$ satisfies the conditions of Theorem ~\ref{thm:ram_order_ind} and Theorem ~\ref{thm:general_generators_relations}. Then, for any spin curve $(\sx,\Delta,L)$ so that $\sx$ and $\sx'$ have the same coarse space $X = X'$ with the same set of ramified points, and $\sx$ has signature $(g;e_1,\ldots, e_r;\delta)$ with $e_1 \leq e_2 \leq \cdots e_r$ so that
$$
e_i \begin{cases}
	= e_i' &\text{ if }i \notin J\\
	\leq e_i' &\text{ if } i \in J
\end{cases}$$
then Theorem ~\ref{thm:general_generators_relations} holds for $(\sx,\Delta,L)$
\end{cor}
\begin{proof}
For $t \leq i \leq r$. Let $(\sx_i,\Delta, L_i)$ be the log spin curve with the same coarse space as $L'$ and with the same stacky points as $(\sx',\Delta,L),$ having signature $(g;e_1, \ldots e_{i-1}, e_i, e_i, \ldots, e_i;\delta),$ and let 
$J_i = \{i,\ldots, r\}.$ Note that $(\sx_k,L_k,J_0) = (\sx',L',J')$ and $(\sx_r ,L_r ,J_r) = (\sx, L, J)$.

Let $(*_i)$ denote the condition that $(\sx_i,\Delta, L_i)$ satisfies the conditions $(\sx',\Delta, L)$ satisfies the conditions of Theorem ~\ref{thm:ram_order_ind} and Theorem ~\ref{thm:general_generators_relations} and $(\sx_i, L_i, J_i)$ is admissible.

Since $(*_0)$ holds by assumption, by induction on $i$, it suffices to show that if $(*_i)$ holds then so does $(*_{i+1}).$ 
Indeed, by applying Lemma ~\ref{lem:admissible_subset}, it follows $(\sx_i, L_i, J_{i+1})$ is admissible, as $J_{i+1} \subset J_i$. Then, applying Theorem ~\ref{thm:ram_order_ind} $\frac{e_{i+1}-e_i}{2}$ times, we see $(*_{i+1})$ holds.
\end{proof}

@@

In this case, we claim $(\sx',L',\{1\})$ is admissible. Indeed, (Ad-
i) holds by part (a) of Lemmas
~\ref{lem:sat-two-induction-high-genus} and ~\ref{lem:deg1-sat-ind},
(Ad-ii) holds by part (b) of Lemmas
~\ref{lem:sat-two-induction-high-genus} and ~\ref{lem:deg1-sat-ind},
and (Ad-iii) holds as $5\deg L \geq 5(g-1)> 2g-1.$ Therefore, by
Lemma ~\ref{lem:n-3s-admissible} , Any triple of the form $(\sx', L',
\{1,\ldots, r\})$ is admissible, with $e_1 = \cdots = e_r = 3$.
Hence, by Corollary ~\ref{cor:raise-stacky-order}, this theorem 
holds for all log spin curves $(\sx, L, J)$. 
\end{proof}

@@

\begin{example}
\label{eg:base- 1-33}
Let $\sx'$ be a stacky curve so that $\halfcan' = P - Q + \frac{1}{3}P_1 + \frac{1}{3}P_2$. In this example, we show
$$R_{\halfcan'} \cong k[u, x_3, y_3, y_4]/(x_3 y_3- \alpha uy_4, y_4^2 - \beta x_3^2 u - \gamma y_3^2u)$$
and that $(\sx',\{3,3\})$ is admissible.

@@

We next check $(\sx',\{1,2\})$ is admissible. We clearly have two generators $x_3, y_3$ in degree 3 with a pole of order $\frac{3- 1}{2}$ by construction. Hence, (Ad-i) holds. We next check (Ad-ii) for the point $P_1$, as the case of $P_2$ is symmetric. Here, by construction
$$
\frac{-\ord_{P_1}(z)}{\ord(z)} = \begin{cases}
	0 &\text{ if }z \in \{u, y_3\}\\
	\frac{1}{4} &\text{ if }z = y_4
\end{cases}.$$
Since $0 < \frac{1}{4} < \frac{1}{3}$, (Ad-ii) holds.
Finally, to check (Ad-iii) note that $\max_{d \geq 0}S_{\sigma,J}(i,d) = 0.$ Therefore, $\deg \lfloor 5L \rfloor  = 2 > 1 = 2 \cdot 1 - 1 + 0$.
\end{example}

@@

Note that $L'$ is admissible, as can be checked in a fashion analogous to the verification in Example ~\ref{eg:base- 1-33}.

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $\max(5, e_1, \ldots, e_r)$ and has minimal relations in degree at most $2e,$ so long as $\sigma$ does not lie in a
finite list of exceptional cases, as listed in Table
~\ref{table:g-1-exceptional}.
\end{thm}

@@

First, we deal with the case $\sx$ has at least 1 stacky point. By Example ~\ref{eg:exception-1-5}, if $L' = P - Q + \frac{2}{5}P_1,$ then $R_{\halfcan'}$ is generated in degrees $2,3,$ and $5$ with a single relation in degree 12, is admissible, and satisfies the hypotheses of Theorem ~\ref{thm:ram_order_ind}. While $L'$ itself is an exceptional case, as it has a generator in degree $12 > 2 \cdot 5$, after applying Theorem ~\ref{thm:ram_order_ind}, we see that $P - Q + \frac{3}{7}P_2$ does satisfy the constraints of this theorem, because only relations in degree $\leq 14 = 2 \cdot 7$ are added, while the previous relation lies in degree $12$, which is less than $2 \cdot 7 = 14$. Therefore, the Theorem holds for $L' = P - Q + \frac{3}{7}P_2$. Then, applying Theorem ~\ref{thm:ram_order_ind} $\frac{e-7}{2}$ times shows that the Theorem holds for $L = P - Q + \frac{e- 1}{2e}P_1$.

Second, we deal with the case that $\sx$ has at least two stacky points. If $(\sx',\Delta,L')$ is a spin canonical curve so that $L' = P - Q + \frac{1}{3}P_1+ \frac{1}{3}P_2$, then, by Example ~\ref{eg:base- 1-33}, it follows that $(\sx', \{3,3\})$ is admissible and satisfies the hypotheses of Theorem ~\ref{thm:ram_order_ind}. Therefore, applying Lemma ~\ref{lem:n-3s-admissible} $r-2$ times, we see that for $(\sx',\Delta',L')$ with $L' = P - Q + \sum_{i = 1}^{r}\frac{1}{3}P_i,$ the pair $(\sx',\{3, \ldots, 3\})$ is admissible. Assume for convenience that $e_1 \leq e_2 \leq \cdots \leq e_r$. Then, by Corollary ~\ref{cor:raise-stacky-order}, this theorem holds for $L' = P - Q + \sum_{i = 1}^{r}\subhalf {e_i}P_i$, as desired.

@@

\begin{lem}
\label{lem:g_0_admissible_cases}
Let $(\sx', \Delta, \halfcan')$ be a log spin curve over a
perfect field $k$, so that $\sx$ has signature $\sigma = (0; e_1,
\ldots, e_r; 0)$. Then $R' := R(\sx', \Delta, \halfcan')$ has a
presentation

is
generated as a $k$-algebra by elements of degree at most $e = \max(5
, e_1, \ldots, e_r)$ with relations in degree at most $2e$ and $(\sx
', J)$ is admissible for the following cases:

@@

We can also always find a presentation for these cases such that
they satisfy (Ad-i) and (Ad-ii) and that $in_\prec(I')$ is
minimally generated by products of two monomials. The procedure for
verifying these are similar to those for the given example.

>>>>

Let $R'$ be the canonical ring associated to $\sx'$. Define $(\sx',L',J)$
to be {\bf admissible} if the $R'$ admits a presentation
\begin{align*}
	R' \cong \left( k[x_1, \ldots, x_m] \otimes k[y_{i, e_i'}]_{i \in J} \right)/I'
\end{align*}

@@

\begin{lem}
\label{lem:admissible_subset}
If $(\sx',\Delta,L')$ is a log spin curve, $(\sx',L', J)$ is admissible and $W \subset J$ is any subset,
then $(\sx', L', W)$ is also admissible.
\end{lem}

@@

\begin{thm}
\label{thm:ram_order_ind}
Suppose $(\sx', \Delta, \halfcan')$ is a log spin curve with coarse
space $X'$ and signature $\sigma = (g; e_1', \ldots, e_r'; \delta)$ and either 
\begin{enumerate}
	\item  $(\sx',\halfcan', J)$ is admissible with generators $x_j, y_{i, e_i'} \in R' = R_{L'}$ with $i \in J$ as in Definition ~\ref{defn:admissible}
	\item or $\sigma$ is one of the signatures described in Remark ~\ref{rem:three-cases}, with $y_{1, 3} = y_{2,3} = \ldots, y_{r,3}$ be a rational section of $\sco(\halfcan_X)$ so that $\ord_{P_i}^{\halfcan_X'}(y_{1,3}) = 1$ at $P_i$ for all $i = 1,\ldots, r$.
\end{enumerate}
Let
$(\sx, \Delta, \halfcan)$ be another log spin curve
with coarse space $X = X'$ and signature $(g; e_1, \ldots, e_r;
\delta)$ such that $e_i = e_i' + 2$ for all $i \in J$ and
$e_j = e_j'$ for $j \notin J$. Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i, e_i} \in
		H^0(\sx, e_i(K_\sx))$ so that
		\begin{align*}
			-\ord_{Q_i}
^{\halfcan_X'}(y_{i, e_i}) = \frac{e_i - 1}{2}
		\end{align*}
		and
		\begin{align*}
			\frac{-\ord_{Q_j}
^{\halfcan_X'}(y_{i, e_i})}{\deg (y_{i, e_i})} \leq 
\subhalf{
			e_j'} - \frac{1}{\deg(y_{i, e_i})e_j'}
		\end{align*}
		for all $j \in J$ with $j \neq i$.
	\item[(b)] If $(\sx',\halfcan',J)$ is admissible, then $y_{i, e_i'}^a y_{i, e_i}^b$ with $a \geq 
0,
		b > 0$ span $R$ over $R'$ and the elements $y_{i, e_i}$ minimally
		generate $R$ over $R'$.
	\item[(c)] If $(\sx',\halfcan',J)$ is admissible, equip $k[y] = k[y_{i, e_i}]_{i \in J}$ and $k[x]$ with
		any graded monomial order and $k[y, x] = k[y] \otimes k[x]$ with
		block order. Let $R = k[y, x]/I.$ Then,
		\begin{align*}
			in_\prec(I) = in_\prec(I')k[x, y] + \langle y_{i, e_i}z \mid z 
			\neq y_{i, e_i}, y_{i, e_i'} \rangle
		\end{align*}
		as $z$ ranges over all generators of $R$.
	\item[(d)] The triple $(\sx, \halfcan', J)$ is admissible.
\end{enumerate}
\end{thm}

\begin{proof}
We will only prove this in the case (1), that $(\sx', \halfcan', J)$ is admissible, as the case (2) that $\sigma$ is as in Remark ~\ref{rem:three-cases} is completely analogous.

@@

Finally, we check (d), by checking $(Ad-i),(Ad-ii)$ and $(Ad-iii)$ are satisfied. We know (Ad-i) holds by part (b), taking the $y_{
i, e_i}$ as the generators in degree $e_i$. Next, (Ad-ii) is
strictly monotonic in the $e_i$ and hence also holds for $(\sx, J)$.
Finally, if (Ad-iii) holds for $e$ then it holds for $e + 2$ by
definition. This is where we use that (Ad-iii) holds for $d > 0$
and not just for $d = 0$.
\end{proof}

\begin{cor}
\label{cor:raise-stacky-order}
Suppose $(\sx',L',J')$ is admissible with signature $\sigma' = (e_1',\ldots,e_r')$ or $\sigma$ satisfies one of the conditions of Remark ~\ref{rem:three cases}. Let $J' = \{t,t+1,\ldots,r\}$ and $e_1' \leq e_2' \leq \cdots \leq e_t' = e_{t+1}' =\cdots = e_r'$
so that $(\sx',\Delta, L)$ satisfies the conditions of Theorem ~\ref{thm:ram_order_ind} and Theorem ~\ref{thm:general_generators_relations}. Then, for any spin curve $(\sx,\Delta,L)$ so that $\sx$ and $\sx'$ have the same coarse space $X = X'$ with the same set of ramified points, and $\sx$ has signature $(g;e_1,\ldots, e_r;\delta)$ with $e_1 \leq e_2 \leq \cdots e_r$ so that
$$
e_i \begin{cases}
	= e_i' &\text{ if }i \notin J\\
	\leq e_i' &\text{ if } i \in J
\end{cases}$$
then Theorem ~\ref{thm:general_generators_relations} holds for $(\sx,\Delta,L)$
\end{cor}
\begin{proof}
For $t \leq i \leq r$. Let $(\sx_i,\Delta, L_i)$ be the log spin curve with the same coarse space as $L'$ and with the same stacky points as $(\sx',\Delta,L),$ having signature $(g; e_1, \ldots. e_{i-1}, e_i, e_i, \ldots, e_i;\delta),$ and let 
$J_i = \{i,\ldots, r\}.$ Note that $(\sx_k,L_k,J_0) = (\sx',L',J')$ and $(\sx_r ,L_r ,J_r) = (\sx, L, J)$.

Let $(*_i)$ denote the condition that $(\sx_i,\Delta, L_i)$ satisfies the conditions $(\sx',\Delta, L)$ satisfies the conditions of Theorem ~\ref{thm:ram_order_ind} and Theorem ~\ref{thm:general_generators_relations} and $(\sx_i, \halfcan_i, J_i)$ is admissible.

Since $(*_0)$ holds by assumption, by induction on $i$, it suffices to show that if $(*_i)$ holds then so does $(*_{i+1}).$ 
Indeed, by applying Lemma ~\ref{lem:admissible_subset}, it follows $(\sx_i, \halfcan_i, J_{i+1})$ is admissible, as $J_{i+1} \subset J_i$. Then, applying Theorem ~\ref{thm:ram_order_ind} $\frac{e_{i+1}-e_i}{2}$ times, we see $(*_{i+1})$ holds.
\end{proof}

@@

In this case, we claim the theorem holds for $(\sx',\Delta, L')$. Indeed, the bound on the generators holds by part (a) of Lemmas
~\ref{lem:sat-two-induction-high-genus} and ~\ref{lem:deg1-sat-ind},
and the bound on the relations holds by part (b) of Lemmas
~\ref{lem:sat-two-induction-high-genus} and ~\ref{lem:deg1-sat-ind}.
Therefore, by Corollary ~\ref{cor:raise-stacky-order}, this theorem 
holds for all log spin curves $(\sx, L, J)$. 
\end{proof}

@@

\begin{example}
\label{eg:base- 1-33}
Let $\sx'$ be a stacky curve so that $\halfcan' = P - Q + \frac{1}{3}P_1 + \frac{1}{3}P_2$. In this example, we show
$$R_{\halfcan'} \cong k[u, x_3, y_3, y_4]/(x_3 y_3- \alpha uy_4, y_4^2 - \beta x_3^2 u - \gamma y_3^2u)$$
and that $(\sx', \halfcan',\{1,2\})$ is admissible.

@@

We next check $(\sx', \halfcan',\{1,2\})$ is admissible. We clearly have two generators $x_3, y_3$ in degree 3 with a pole of order $\frac{3- 1}{2}$ by construction. Hence, (Ad-i) holds. We next check (Ad-ii) for the point $P_1$, as the case of $P_2$ is symmetric. Here, by construction
$$
\frac{-\ord_{P_1}(z)}{\ord(z)} = \begin{cases}
	0 &\text{ if }z \in \{u, y_3\}\\
	\frac{1}{4} &\text{ if }z = y_4
\end{cases}.$$
Since $0 < \frac{1}{4} < \frac{1}{3}$, (Ad-ii) holds.
Finally, to check (Ad-iii) note that $\max_{d \geq 0}S_{\sigma,J}(i,d) = 0.$ Therefore, $\deg \lfloor 5L \rfloor  = 2 > 1 = 2 \cdot 1 - 1 + 0$.
\end{example}

@@

Note that $(\sx', \halfcan', \{1\})$ is admissible, as can be checked in a fashion analogous to the verification in Example ~\ref{eg:base- 1-33}.

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $\max(5, e_1, \ldots, e_r)$ and has relations in degree at most $2e,$ so long as $\sigma$ does not lie in a
finite list of exceptional cases, as listed in Table
~\ref{table:g-1-exceptional}.
\end{thm}

@@

First, we deal with the case $\sx$ has at least 1 stacky point. By Example ~\ref{eg:exception-1-5}, if $L' = P - Q + \frac{2}{5}P_1,$ then $R_{\halfcan'}$ is generated in degrees $2,3,$ and $5$ with a single relation in degree 12, $(\sx, \halfcan', \{1\})$ is admissible, and satisfies the hypotheses of Theorem ~\ref{thm:ram_order_ind}. While $L'$ itself is an exceptional case, as it has a generator in degree $12 > 2 \cdot 5$, after applying Theorem ~\ref{thm:ram_order_ind}, we see that $P - Q + \frac{3}{7}P_2$ does satisfy the constraints of this theorem, because only relations in degree $\leq 14 = 2 \cdot 7$ are added, while the previous relation lies in degree $12$, which is less than $2 \cdot 7 = 14$. Therefore, the Theorem holds for $L' = P - Q + \frac{3}{7}P_2$. Then, applying Theorem ~\ref{thm:ram_order_ind} $\frac{e-7}{2}$ times shows that the Theorem holds for $L = P - Q + \frac{e- 1}{2e}P_1$.

Second, we deal with the case that $\sx$ has at least two stacky points. If $(\sx',\Delta,L')$ is a spin canonical curve so that $L' = P - Q + \frac{1}{3}P_1+ \frac{1}{3}P_2$, then, by Example ~\ref{eg:base- 1-33}, it follows that the signature $\sigma = (1;3, 3; \delta)$ satisfies the hypotheses of Theorem ~\ref{thm:ram_order_ind}. Therefore, applying Lemma ~\ref{lem:n-3s-admissible} $r-2$ times, we see that for $(\sx',\Delta',L')$ with $L' = P - Q + \sum_{i = 1}^{r}\frac{1}{3}P_i,$ the pair $(\sx', \halfcan', \{1, \ldots, r\})$. Assume for convenience  $e_1 \leq e_2 \leq \cdots \leq e_r$. Then, by Corollary ~\ref{cor:raise-stacky-order}, this theorem holds for $L' = P - Q + \sum_{i = 1}^{r}\subhalf {e_i}P_i$, as desired.

@@

\begin{lem}
\label{lem:g_0_admissible_cases}
Let $(\sx', \Delta, \halfcan')$ be a log spin curve over a
perfect field $k$, so that $\sx$ has signature $\sigma = (0; e_1,
\ldots, e_r; 0)$. Then, $R' := R(\sx', \Delta, \halfcan')$ has a
presentation generated as a $k$-algebra by elements of degree at most $e = \max(5
, e_1, \ldots, e_r)$ with relations in degree at most $2e$ and $(\sx
', \halfcan', J)$ is admissible for the following cases:

@@

We can also always find a presentation for these cases such that
they satisfy (Ad-i) and (Ad-ii) and that $in_\prec(I')$ is
generated by products of two monomials. The procedure for
verifying these are similar to those for the given example.
